---
title: "Item Response Theory (IRT) Implementation and Application"
output: html_document
date: "2025-04-28"
---
Part 1. Classical Test Theory
The dataset "data.sim.rasch" simulates binary item responses from a Rasch model. 
這段程式碼用於載入一個模擬的 Rasch 模型測驗作答資料，

Step 1.Examining the structure and content of the dataset to understand its dimensionality (2000 persons × 40 items) and verify its suitability for psychometric analysis.
總共有 2000 位受試者對40題二元題目的作答結果(1=作答正確，0=作答錯誤)。
```{r}
library(TAM)   #讀入 TAM 套件
data(data.sim.rasch) #從 TAM 套件中抓出名為 data.sim.rasch 的資料集
head(data.sim.rasch) #列出該資料集頭幾行的資料
str(data.sim.rasch) #共有2000筆
```

Step 2.Computing examinees' raw scores (total correct responses) and item mean scores (item easiness indices) as part of classical test theory preprocessing. 
計算整體表現指標：受試者的總分和每題的易度(即通過率)
```{r}
resp <- data.sim.rasch #複製data.sim.rasch資料，並用resp作為新檔名
scores <- apply(resp,1,sum) #計算每位受試者的總分(作答正確的題數)
head(scores) #列出頭幾行的得分資料
itemscores <- apply(resp,2,mean) #計算每題的平均得分(正答率)
head(itemscores)
```

Step 3.Discrimination題目鑑別度
```{r}
cor(scores,resp[,1]) #第一題的鑑別度=第一題與總分的相關
```

Step 4.Computing item-total correlations (discrimination indices), adjusting for item overlap by excluding the target item from total scores.
計算每一題與「去除該題後的總分」之間的相關係數，作為該題的「鑑別度」指標
```{r}
cor(scores-resp[,1],resp[,1]) #更精確的鑑別度(避免項目與自己相關)
```


```{r}
#計算所有題目的鑑別度
disc <- apply(resp,2,function(x){cor(scores-x,x)})
disc
```

#Another Method: CTT
```{r}
library(CTT)
resp <- data.sim.rasch
IA <- itemAnalysis(resp) #將題目分析結果存入 IA 這一個物件
IA$itemReport
IA$alpha #Cronbach's alpha = 0.90
```


```{r}
data(CTTdata) #資料集
str(CTTdata) ##資料集:100位樣本、20題題目
data(CTTkey) #正確答案
head(CTTdata)
CTTkey
```

```{r}
resp <- score(CTTdata,CTTkey,output.scored = TRUE)
str(resp)
resp$scored
```

```{r}
scored <- resp$scored   # 取出每題得分（0/1）
itemAnalysis(scored) #Cronbach's alpha = .83，代表整份測驗有良好的信度
total_score <- resp$score  # 取出總分

# 計算未修正的鑑別度(item-total correlation)
item_total_correlation <- cor(scored, total_score)
round(item_total_correlation,2)

# 計算修正後的鑑別度（corrected item-total correlation）
disc <- apply(scored, 2, function(x) cor(total_score - x, x))
round(disc,2)
```

Step 5. Drawing Diagram
```{r}
item_names <- colnames(scored)
plot_df <- data.frame(
  item = factor(item_names, levels = item_names),
  uncorrected = as.vector(item_total_correlation),
  corrected = disc
)

library(ggplot2)

ggplot(plot_df, aes(x = item)) +
  geom_line(aes(y = uncorrected, group = 1, color = "Original"), size = 0.5) +
  geom_point(aes(y = uncorrected, color = "Original"), size = 1) +
  geom_line(aes(y = corrected, group = 1, color = "Corrected"), size = 0.5, linetype = "dashed") +
  geom_point(aes(y = corrected, color = "Corrected"), shape = 15, size = 1) +
  geom_hline(yintercept = 0.3, linetype = "dotted", color = "gray40") +
  labs(
    title = "Item Discrimination Comparison Chart",
    x = "Item Number",
    y = "Discrimination (Pearson Correlation)",
    color = "Type"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("Original" = "blue", "Corrected" = "red")) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

Part 2.Item Response Theory (IRT)
Step 1.Computing the correct rate in Rasch
```{r}
delta <- 0.6  # 題目難度
theta <- 1.0  # 學生能力
prob <- exp(theta - delta) / (1 + exp(theta - delta))
prob
```

Step 2.Draw ICC
```{r}
delta <- 1
theta <- seq(-3, 3, 0.01)
prob <- exp(theta - delta) / (1 + exp(theta - delta))
plot(theta, prob, type = "l",
     xlab = expression(theta),
     ylab = "Probability of Correct Response",
     main = paste("ICC for Rasch Model (Difficulty =", delta, ")"))
```

Step 3.Draw different difficulty of ICC
```{r}
delta1 <- -1
delta2 <- 0
delta3 <- 1
prob1 <- exp(theta - delta1) / (1 + exp(theta - delta1))
prob2 <- exp(theta - delta2) / (1 + exp(theta - delta2))
prob3 <- exp(theta - delta3) / (1 + exp(theta - delta3))

plot(theta, prob1, type = "l", col = "blue", lwd = 2,
     xlab = expression(theta), ylab = "Probability of Correct Response",
     main = "Item Characteristic Curves (ICC)")
lines(theta, prob2, col = "red", lwd = 2)
lines(theta, prob3, col = "green", lwd = 2)
legend("bottomright", legend = c(expression(delta == -1), expression(delta == 0), expression(delta == 1)),
       col = c("blue", "red", "green"), lwd = 2)
```

Step 4.Item Characteristic Curve (ICC) of the Two-Parameter Logistic (2PL) IRT Model
雙參數 IRT 模型 (2PL) 的 ICC
```{r}
delta <- 0
a1 <- 0.5
a2 <- 1.0
a3 <- 1.5
prob1 <- 1 / (1 + exp(-a1 * (theta - delta)))
prob2 <- 1 / (1 + exp(-a2 * (theta - delta)))
prob3 <- 1 / (1 + exp(-a3 * (theta - delta)))

plot(theta, prob1, type = "l", col = "blue", lwd = 2,
     xlab = expression(theta), ylab = "Probability of Correct Response",
     main = "Item Characteristic Curves with Different Discrimination (a)")
lines(theta, prob2, col = "red", lwd = 2)
lines(theta, prob3, col = "darkgreen", lwd = 2)
legend("bottomright", legend = c(expression(a == 0.5), expression(a == 1.0), expression(a == 1.5)),
       col = c("blue", "red", "darkgreen"), lwd = 2)
```

Step 5.Plotting ICCs for 20 Randomly Generated Parameters
```{r}
a <- runif(20, 0.5, 1.5)
b <- runif(20, -2, 2)
colors <- rainbow(20)

plot(theta, 1 / (1 + exp(-a[1] * (theta - b[1]))), type = "l", col = colors[1], lwd = 2,
     xlab = expression(theta), ylab = "Probability",
     main = "ICC Curves for 20 Items")
for(i in 2:20){
  prob <- 1 / (1 + exp(-a[i] * (theta - b[i])))
  lines(theta, prob, col = colors[i], lwd = 2)
}
legend("bottomright", legend = paste("Item", 1:20), col = colors, lwd = 2, cex = 0.6)
```

Step 6.Analyzing data.sim.rasch Using the TAM Package
使用 TAM 套件分析 data.sim.rasch
```{r}
library(TAM)
data("data.sim.rasch")
head(data.sim.rasch)
```
Step 7.Estimating Rasch Model Parameters Using Joint Maximum Likelihood Estimation (JMLE)
聯合最大概似估計法 (JMLE) 估計 Rasch 模型參數
```{r}
mod1 <- tam.jml(data.sim.rasch)
summary(mod1)
mod1$item
mod1$WLE
mod1$WLEreliability

summary(mod1$item)
summary(mod1$WLE)
hist(mod1$WLE, main = "Distribution of Estimated Abilities (WLE)", xlab = "Ability (WLE)")
```

Step 8.Plotting the ICC for Each Item (Saved as a PDF)
繪製每個項目的 ICC (儲存為 PDF)
```{r}
pdf("ICCplots.pdf")
plot(mod1)
dev.off()

for (i in 1:mod1$nitems){
  plot(mod1, items = i)
}
```

Step 9.Plotting Wright Map
繪製 Wright Map
```{r}
library(WrightMap)
wrightMap(mod1$WLE, mod1$xsi, item.side = itemClassic, main = "Wright Map (Classical Item Side)")
wrightMap(mod1$WLE, mod1$xsi, item.side = itemModern, person.side = personHist, main = "Wright Map (Modern Item & Person Side)")

```

Part 3.Application of the Generalized Partial Credit Model (GPCM) to Polytomous Data
廣義部分計分模型 (GPCM) 在多級評分資料上的應用

Step 1. Loading Packages
載入套件
```{r}
library(psych)
library(mirt)
```

Step 2.Loading Big Five Dataset
載入 Big Five 資料集
```{r}
library(mirt)
library(psych)
data(bfi)
head(bfi)
str(bfi)

# 選取前 25 個題目
IRTgrm.items <- bfi[, 1:25]
```

Step 3.Fitting a Unidimensional GPCM Model
擬合單維度的 GPCM 模型
```{r}
IRTgrm3 <- mirt(data = IRTgrm.items, model = 1, itemtype = "graded")
summary(IRTgrm3)
coef(IRTgrm3, IRTpars = TRUE)
```

Step 4.Plotting the Option Characteristic Curves (OCCs) and Item Information Functions for the First and Fifth Items
繪製第一個和第五個項目的選項特徵曲線和項目資訊函數
```{r}
itemplot(IRTgrm3, 1, type = "trace")
itemplot(IRTgrm3, 1, type = "info")
itemplot(IRTgrm3, 5, type = "trace")
itemplot(IRTgrm3, 5, type = "info")
```

Step 5.Plotting the Option Characteristic Curves and Item Information Functions for All Items
繪製所有項目的選項特徵曲線和項目資訊函數
```{r}
plot(IRTgrm3, type = "trace", main = "Item Trace Lines (GPCM)")
plot(IRTgrm3, type = "info", main = "Item Information Functions (GPCM)")
```

